---
title: "Code Portfolio"
subtitle: "Data Structure and Transformation"
author: "Cho-Lin Chiang"
date: "2019/4/16"
output:
  html_document:
    number_sections: true
    code_folding: show
    theme: lumen
    toc: yes
    toc_float: TRUE
    toc_depth: 2
---

<br> 

# Introduction

Week 5 - In this class we went over the data structures of matrices and dataframe, and looked more indepth into methods of dealing with missing data.

<br> 

# Types of missing data

There is, in general, 4 different types of missing data. <br>
<br>

1. Structurally Missing Data - is data that is missing because it doesn't exist in the first place. Take for example, a survey that collects data on smokers. If a survey respondent is not a smoker, data under "years of smoking" will be left blank since he/she doesn't smoke. To deal with this type of data, one can exclude these observations with structurally missing data.
<br>

2. Missing at Random (MAR) - is data that is missing randomly. There are ways to predict the missing data, such as using advanced imputation method like *multiple inputation* with `library(Hmisc)` and `library(missForest)`. 
   i) `impute()` function fills missing value with with statistical methods such as mean, max, or medium. The default is median. e.g `with(data, impute(x, mean))`
    ii) `aregImpute()` function fills missing value with additive regression, bootstrapping, and predictive mean matching. e.g. `aregImpute(~ a + b + c + d, data=data, n.impute=4)`
<br>

3. Missing Not at Random (MNAR) - is data where imputation cannot be used, largely due to the uncertainty towards probabilistic value of the missing values.
<br> 

4. Missing Completely at Randm (MCAR) - is data points being a random subset of the overall dataset.
    i) Listwise deletion - using `na.omit(data)`
    ii) Variable deletion - if more than 60% of the data is missing and this particular variable is not significant, observations can be omitted. 
    iii) Pairwise deletion - preferred over listwise deletion due to the limited loss of data. It maximizes all data available. 

# Data Structure Practice
Chapter 8.7  Phillips, N. D. (2016). "Yarrr! The pirate's guide to R" https://bookdown.org/ndphillips/YaRrr/test-your-r-might-pirates-and-superheroes.html

<br>
1. Combine the data into a single dataframe. Complete all the following exercises from the dataframe!
```{r, warning = FALSE, message=FALSE, results="hide"}
survey <- data.frame("name" = c("Astrid","Lea","Sarina","Remon","Letizia","Babice","Jonas","Wendy","Niveditha","Gioia"),
                     "sex" = c("F","F","F","M","F","F","M","F","F","F"),
                     "age" = c(30,25,25,29,22,22,35,19,32,21),
                     "superhero"=c("Batman","Superman","Batman","Spiderman","Batman","Antman","Batman","Superman","Maggott","Superman"),
                     "tattos" = c(11, 15, 12, 5, 65, 3, 9, 13, 900, 0))
```

<br>

2. What is the median age of the 10 pirates?
```{r, warning = FALSE,message=TRUE}
median(survey$age)
```

<br>

3. What was the mean age of female and male pirates separately?
```{r, warning = FALSE,message=TRUE}
mean(survey$age[survey$sex=="F"])
mean(survey$age[survey$sex=="M"])
```

<br>

4. What was the most number of tattoos owned by a male pirate?
```{r, warning = FALSE,message=TRUE}
max(survey$tattos[survey$sex=="M"])
```

<br>

5. What percent of pirates under the age of 32 were female?
```{r, warning = FALSE, message=TRUE}
young <- subset(x=survey, 
                subset = age<32)

table(young$sex)/sum(table(young$sex))
```

<br>

6. What percent of female pirates are under the age of 32?
```{r, warning = FALSE, message=TRUE}
female <- subset(x=survey,
                 subset = sex=="F")

table(female$age<32)/sum(table(female$age<32))
```

<br>

7. Add a new column to the dataframe called tattoos.per.year which shows how many tattoos each pirate has for each year in their life.
```{r, warning = FALSE, message=TRUE}
tattoos.per.year <- survey$tattos/survey$age
newsurvey <-cbind(survey,tattoos.per.year)
newsurvey
```

<br>

8. Which pirate had the most number of tattoos per year?
```{r, warning = FALSE, message=TRUE}
x <- which.max(newsurvey$tattoos.per.year)
newsurvey[x,]
```

<br>
