---
title: "Code Portfolio"
subtitle: "Clustering - Hierarchical and K-Means"
author: "Cho-Lin Chiang"
date: "2019/5/28"
output:
  html_document:
    number_sections: true
    code_folding: show
    theme: lumen
    toc: yes
    toc_float: TRUE
    toc_depth: 2
---

<br> 

# Introduction
Week 11 - This class goes over the types of clustering and calculation methods in R, with a focus on K-means and Hierarchical clusterings.

<br> 

# Types of Clustering


1. **Partitional Clustering** - divides data into non-overlapping subsets.

2. **Hierarchical Clustering** - divides data into nested clusters organized as a hierarchical tree.

<br>

## K Means Clustering

K Means Clustering is a method that initially chooses a number of clusters by random, and reiterate to improve the clusters. The main goal of defining clusters is to minimize intra-cluster variation.

Steps:

1. Randomly choose number of clusters.

2. Based on (1), select the initial cluster centroid as starting points.

3. Assign each data point to the cluster centroid (2) it is closest to.

4. Recalculate centroids.

5. Reassign clusters.

6. Repeat (4) and (5)

<br>

## Echo Clustering

Using package `factoextra` and its function `fviz_nbclust()`, stating `method = "wss"`, plot graph that shows number of clusters to total sum of squares. Find the optimal number of clusters that looks like an elbow on the graph.

<br>

## Silhouette Method

Similar to Echo clustering, use package `factoextra` and its function `fviz_nbclust()`, stating `method = "silhouette"`, plot graph that shows number of clusters to total sum of squares. Find the number where a spike occurs.(However, this method seems to always suggest 2 as the optimal clustering number)

<br>

## Gap Statistic Method

Using the gap function, such as `fviz_nbclust(clusters, kmeans, method = "gap_stat")+ theme_classic()`, identify the best number of clusters from the plot. 



<br>

# Calculating K-means in R

We can use the function `kmeans()` to calculate k-means. 

```{r, warning = FALSE, message=FALSE, results="hide"}
library(tidyverse)
library(cluster)
library(factoextra)
```

```{r, warning = FALSE,message=FALSE}
df <- scale(na.omit(USArrests))
k1 <- kmeans(df, centers = 2, nstart = 25)
str(k1)
```

<br>

We can visually plot the results with `fviz_cluster()`.
 
```{r, warning = FALSE,message=FALSE}
fviz_cluster(k1, data = df)
```

<br>
 