---
title: "Code Portfolio"
subtitle: "Plotting Systems and Graphical Device"
author: "Cho-Lin Chiang"
date: "2019/5/14"
output:
  html_document:
    number_sections: true
    code_folding: show
    theme: lumen
    toc: yes
    toc_float: TRUE
    toc_depth: 2
---

<br> 

# Introduction

Week 9 - Executive session - covered topics concerning the different plotting systems in R, as well as graphics devices.

<br> 

# Plotting Systems

## Base Plotting System

This plotting system comes imbedded within the original R. The defining characteristic is that plots are built from scratch, command by command. It is a cumulative process that makes going back to previous step impossible once a new command is incorporated. 


```{r, warning = FALSE,message=FALSE}
hist(cars$speed)
```

<br>

```{r, warning = FALSE, message=FALSE, echo=FALSE}
with(cars, plot(speed, dist)) +
title(main = "Speed and Distance")
```

<br>

## Lattice System

The lattice system requires `lattice` package installation, and is most convenient when panel plots are needed (effects of *y* on *x* across levels of *z*). However, it is considered anti-intuitive because it is difficult to annotate and cannot add to plot after it has been created)

```{r, warning = FALSE,message=FALSE}
library(lattice)
library(nycflights13)
xyplot(arr_delay~ air_time | month, data=flights, layout=c(4,3))
```

<br>

# Graphics Devices

Graphics devices are devices that make plots appear. And, plotting graphs in R, you'd come across the different devices:

1. window (type of device: screen)

2. PDF (type of device: file)

3. PNG/ JPEG (type of device: file)

4. Scalable Vector Graphics (SVG) (type of device: file)



```{r, warning = FALSE,message=FALSE}

```

<br>

```{r, warning = FALSE,message=FALSE}

```

<br>


