---
title: "Code Portfolio"
subtitle: "Data Visualization"
author: "Cho-Lin Chiang"
date: "2019/4/30"
output:
  html_document:
    number_sections: true
    code_folding: show
    theme: lumen
    toc: yes
    toc_float: TRUE
    toc_depth: 2
---

<br> 

# Introduction
Week 7 - Data Visualization's role within Exploratory Data Analysis. 

<br> 

# Basic Concepts
## Measurement Scales
There are 4 levels of measurement scales: 

1. Nominal

2. Ordinal

3. Interval

4. Ratio

<br> 

## Visualization Plots
Examples of visualization plots:

1. **Bar Chart** - nominal scale. Usually x-axis is categorical data (discrete) and y-axis is numerical data (counts, proportion, percentage).

2. **Frequency Histogram** - sensitive to bin placements. As a rule of thumb, 
    i) If `n>150`, you'll need 100 bins
    ii) If `30<n<150`, experiment with different bin numbers.
    iii) If `n<30`, you should consider using graphs other than a frequency histogram.

3. **Box Plot** - relates information such as range, quartiles, and median.

<br>

## Distribution Shapes
Distribution shapes of frequency histogram can take a few forms:

1. Constant

2. Bell-shaped (normal distribution)

3. U-shaped (bimodal)

4. Constantly increasing / decreasing (linear distribution)

5. Exponentially increasing (non-linear distribution)

<br> 

# Visualization in R
The most common and versatile package used in R for data visualization is `ggplot2`.

<br>

# Practice: Chapter 5 in "R for Data Science"

Using the `mpg` dataset as practice, we can create scatterplots with `geom_point`

<br>

```{r, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(ggplot2)
library(tidyverse)
```

```{r, warning = FALSE,message=TRUE}
ggplot(data=mpg) + geom_point(aes(x=displ, y=hwy))
```
This plot shows that `hwy` (highway miles per gallon) trends lower when `displ` (engine displacement) increases. 

<br>

There are plots that are less useful, such as this one comparing `class` and `drv` 
```{r, warning = FALSE,message=TRUE}
ggplot(data=mpg) + geom_point(aes(x=class, y=drv))
```
When we plot categorical variable to categorical variable, it doesn't tell you anything informative. There's no information of whether more midsize vehicles use front-wheel drive, or whether there is any trends since they are nominal data.

<br>

We can convey more information using **color**. In the following plot, `displ` is compared to `hwy`, with colors adapted to `cyl` (number of cylinders).

```{r, warning = FALSE,message=TRUE}
ggplot(data=mpg) + geom_point(aes(x=displ, y=hwy, color = cyl))
```

We can easily tell from this plot that higher `displ` usually means higher `cyl`. 

<br>

Other features such as `size=` and `alpha=` are helpful in determining frequency

```{r, warning = FALSE,message=TRUE}
ggplot(data=mpg) + geom_point(aes(x=displ, y=hwy, size = cyl))
ggplot(data=mpg) + geom_point(aes(x=displ, y=hwy, alpha = cyl))
```

<br>

`shape=`, on the other hand, is useful in differentiating categorical variables, such as `class`.

```{r, warning = FALSE,message=TRUE}
ggplot(data=mpg) + geom_point(aes(x=displ, y=hwy, shape = class))
```

<br>
Trendlines can also be added to convey a clearer message

```{r, warning = FALSE,message=FALSE}
ggplot(data=mpg) + 
  geom_point(aes(x=displ, y=hwy, color = cyl)) +
  geom_smooth(aes(x=displ, y=hwy))
  
```

<br>

