---
title: "Code Portfolio"
subtitle: "Practical Stats and Data Transformation"
author: "Cho-Lin Chiang"
date: "2019/4/23"
output:
  html_document:
    number_sections: true
    code_folding: show
    theme: lumen
    toc: yes
    toc_float: TRUE
    toc_depth: 2
---

<br> 

# Introduction
Week 6 - In this class, we went over key terminologies related to statistics, their definitions, and R functions.

<br> 

# Key Terminology and definition
- **Mean** - average of a group of numbers
- **Median** - the value in the middle of a sorted group of numbers.Can be calculated with `mad()`.
- **Weighted mean** - sum of all values multiplied by weight, and then divided by the sum of weights. We can use the function `weighted.mean(x, w=weight)` under the package `matrixStats`.
- **50th percentile** - similar to median
- **Deviation** - difference between observed value and the mean
- **Variance** - square of standard deviation
- **Standard deviation** - calculated difference of deviation for a group of numbers. Can be calculated with `sd()`.

<br> 

# Practice: Chapter 5 of "R for Data Science"
Exercise 5.2.4 Based on `nycflights13`

<br>
```{r, warning = FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(nycflights13)
```

1. Find all flights that
    i) Had an arrival delay of two or more hours

```{r, warning = FALSE, message=TRUE}
twohourdelay <- filter(flights, arr_delay >= 120)
```
    
    ii) Flew to Houston (`IAH` or `HOU`) 
    
```{r, warning = FALSE, message=TRUE}
houston <- filter(flights, dest == "IAH" | dest == "HOU")
```
    
    iii) Were operated by United, American, or Delta
    
```{r, warning = FALSE, message=TRUE}
carrier <- filter(flights, carrier == "UA" | carrier == "AA" | carrier == "DL")
```
    
    iv) Departed in summer (July, August, and September)
    
```{r, warning = FALSE, message=TRUE}
summer <- filter(flights, month %in% c(7,8,9))
```
    
    v) Arrived more than two hours late, but didn¡¦t leave late 
    
```{r, warning = FALSE, message=TRUE}
arrivelate <- filter(flights, arr_delay>=120 & dep_delay <=0)
```
   
    vi) Were delayed by at least an hour, but made up over 30 minutes in flight 
    
```{r, warning = FALSE, message=TRUE}
madeup <- filter(flights, dep_delay >=60 & arr_delay <30)
```
  
    vii) Departed between midnight and 6am (inclusive)
    
```{r, warning = FALSE, message=TRUE}
midnight <- filter(flights, !(dep_time > 600))
midnight2 <- filter(flights, dep_time >= 000 & dep_time <= 600)
```
   
    viii) Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?
    
```{r, warning = FALSE, message=TRUE}
# Between function is a shortcut of the logical functions used previously. It simplifies the command to such as : 
midnight3 <- filter(flights, between(flights$dep_time, 000, 600))
```

2. How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?

```{r, warning = FALSE, message=TRUE}
# Number of flights with missing dep_time
sum(is.na(flights$dep_time))

# Variables that contain missing data
colSums(is.na(flights))

# These rows may represent flights that were cancelled even though they were scheduled (in the system).
```


<br>

