---
title: "Code Portfolio"
subtitle: "Week 4 - Data Transformation"
author: "Cho-Lin Chiang"
date: "2019/6/4"
output:
  html_document:
    code_folding: show
    theme: lumen
    toc: yes
    toc_depth: 2
---

```{r, warning = FALSE, message=FALSE, results="hide"}

```

<br> 

## Introduction
Data cleaning and preparation take up 80% of time spent in data analysis. And tidyverse workflow suggests a framework to structuring data before any analysis takes place, that includes transforming the data with `dplyr`.

<br> 

## Practice: Chapter 11 of "R for Data Science"

The packages needed in this exercise includes `tidyverse` and `dplyr`
```{r, warning = FALSE, message=FALSE, results="hide"}
library(tidyverse)
library(dplyr)
```

<br>

The function `parse_datetime()` organizes an ISO8601 formatted time from the biggest to smallest:
```{r, warning = FALSE,message=TRUE}
parse_datetime("2010-11-30T2010")
```

<br>

### Exercise 11.3.5.7
Generate the correct format string to parse each of the following dates and times:
```{r, warning = FALSE}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```

<br>

The percentage sign `%` followed by a character defines the format of the date or time stamp.<br>
For Example: <br>
`%B` indicates month spelt in full, such as "January", "February", and "March" <br>
`%b` indicates month shortened, such as "Jan", "Feb", and "Mar" 


```{r, warning = FALSE}
parse_datetime(d1, "%B %d, %Y")
parse_datetime(d2, "%Y-%b-%d")
parse_datetime(d3, "%d-%b-%Y")
parse_datetime(d4, "%B %d (%Y)")
parse_datetime(d5, "%m/%d/%y")
parse_time(t1,"%H %M")
parse_time(t2)
```

<br>
